Storm surge modelling in the Strait of Georgia

### 2014-12-03 RoughWork/manuscript.tex
Added environment canada references

At the ocean surface, meteorological conditions and fresh water input from rivers are included. The ocean surface is forced with momentum and heat fluxes from a 33-km global atmospheric reforecasting model suitable for use in ocean modelling \citep{smith2014new}. Forecasts from the period of 2002-2012 are available. Additionally, the inverse barometric effect of the atmospheric pressure is included, an important consideration in storm surge modelling. One simulation employs the western Canada component of the High Resolution Deterministic Prediction System (HRDPS), a nested 2.5 km resolution atmospheric model provided by Environment Canada. %citation?

@@

\subsection{Strong Wind Event}
Next, a relatively strong wind event recorded at Point Atkinson on Nov 19, 2009 is considered. As recorded in the Environment Canada Historical Climate Data archives, the maximum hourly wind speed at Point Atkinson was 18.6 m s$^{-1}$ and the largest gust observed was 25.8 m s$^{-1}$ or 93 km hr$^{-1}$. In addition, large wind speeds over 15 m s$^{-1}$ persisted for over 9 hours. Yet, the surge associated with this event is rather small when compared to the previous two cases as shown in Figure \ref{fig:nov2009} where we have plotted the wind speed at Point Atkinson, sea surface height anomaly forcing at the open boundaries, and observed and modelled residuals at Point Atkinson. Two significant surge events are noted, the first occurring on Nov 19 with an observed amplitude of about 50 cm at Point Atkinson, and a second, larger surge of 62.5 cm on Nov 20. Surprisingly, the strong wind event of Nov 19 was associated with the smaller of the two surges. Neither of the surges occurred during a high tide and so no high water level events were recorded. Although not shown here, the other three locations also observed a maximum surge on Nov 20, which appears to be linked to a secondary high wind event and another surge entering the system at Tofino and Port Hardy. 
%citation for Environment Canada data?

@@

\begin{figure}
\centering
\includegraphics[scale=0.6]{Figures/nov2009.pdf}
\caption{Wind and storm surge conditions at Point Atkinson for a strong wind event on Nov 19, 2009. Top: Wind speed from Environment Canada observations at Point Atkinson (solid) and from the atmospheric model (CGRF) at a grid point near Point Atkinson (dashed). Middle: Sea surface height anomaly at Tofino (solid) and Port Hardy (dashed) used as forcing conditions at the western and northern open boundaries. Bottom: Observed (solid) and modelled (dashed) residuals at Point Atkinson.  }
\label{fig:nov2009}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[scale=0.6]{Figures/dec2012_weather.pdf}
\caption{Comparison of observed and modelled winds during Dec 2012 storm surge simulation at Point Atkinson (upper) and Sandheads (lower). Observed winds shown in solid black, low resolution CGRF model in dashed black and high resolution model in grey. }
\label{fig:dec2012_weather}
\end{figure}
%don't include direction?

>>>>

At the ocean surface, meteorological conditions and fresh water input from rivers are included. The ocean surface is forced with momentum and heat fluxes from a 33-km global atmospheric reforecasting model suitable for use in ocean modelling \citep{smith2014new}. Forecasts from the period of 2002-2012 are available. Additionally, the inverse barometric effect of the atmospheric pressure is included, an important consideration in storm surge modelling. One simulation employs the western Canada component of the High Resolution Deterministic Prediction System (HRDPS), a nested 2.5 km resolution atmospheric model provided by Environment Canada \citep{ECModel}.

@@

\subsection{Strong Wind Event}
Next, a relatively strong wind event recorded at Point Atkinson on Nov 19, 2009 is considered. As recorded in the Environment Canada Climate Data archives \citep{ECClimateArchive}, the maximum hourly wind speed at Point Atkinson was 18.6 m s$^{-1}$ and the largest gust observed was 25.8 m s$^{-1}$ or 93 km hr$^{-1}$. In addition, large wind speeds over 15 m s$^{-1}$ persisted for over 9 hours. Yet, the surge associated with this event is rather small when compared to the previous two cases as shown in Figure \ref{fig:nov2009} where we have plotted the wind speed at Point Atkinson, sea surface height anomaly forcing at the open boundaries, and observed and modelled residuals at Point Atkinson. Two significant surge events are noted, the first occurring on Nov 19 with an observed amplitude of about 50 cm at Point Atkinson, and a second, larger surge of 62.5 cm on Nov 20. Surprisingly, the strong wind event of Nov 19 was associated with the smaller of the two surges. Neither of the surges occurred during a high tide and so no high water level events were recorded. Although not shown here, the other three locations also observed a maximum surge on Nov 20, which appears to be linked to a secondary high wind event and another surge entering the system at Tofino and Port Hardy.

@@

\begin{figure}
\centering
\includegraphics[scale=0.6]{Figures/nov2009.pdf}
\caption{Wind and storm surge conditions at Point Atkinson for a strong wind event on Nov 19, 2009. Top: Wind speed from Environment Canada observations \citep{ECClimateArchive} at Point Atkinson (solid) and from the atmospheric model (CGRF) at a grid point near Point Atkinson (dashed). Middle: Sea surface height anomaly at Tofino (solid) and Port Hardy (dashed) used as forcing conditions at the western and northern open boundaries. Bottom: Observed (solid) and modelled (dashed) residuals at Point Atkinson.  }
\label{fig:nov2009}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[scale=0.6]{Figures/dec2012_weather.pdf}
\caption{Comparison of observed and modelled winds during Dec 2012 storm surge simulation at Point Atkinson (upper) and Sandheads (lower). Observed winds shown in solid black, low resolution CGRF model in dashed black and high resolution model in grey. Observations are from the Environment Canada Climate Data archives \citep{ECClimateArchive}.}
\label{fig:dec2012_weather}
\end{figure}
%don't include direction?
